MIRSVM: Multi-Instance Support Vector Machine with Bag Representatives

### 2017-09-22 ArticleRevision/article/article.tex
Updated the article for Pattern Recognition to be without markings

\section{Introduction}\label{intro}
In traditional classification, learning algorithms attempt to correctly label unknown samples by finding patterns between training samples and their class label. \textit{Multi-instance} (MI) learning is a generalization of \textit{supervised learning} that has been recently been gaining interest because of its applicability to many real-world problems such as text categorization \citep{Andrews2002}, image classification and annotation \citep{Herman2008,qi2007}, human action recognition \citep{Yi2016}, and drug activity prediction \citep{Dietterich1997}.

@@

Support vector machines represent a set of supervised, linear and nonlinear, classification and regression methods that have theoretical foundations on Vapnik-Chervonenkis (VC) theory~\citep{Huang2006,Schoelkopf2002}. SVM models are similar to other machine learning algorithms and techniques, but research has shown that they usually outperform them in terms of computational efficiency, scalability, and robustness against outliers~\citep{Kecman2016,Melki2016}, which makes them a useful data mining tool for various real-world applications~\citep{chen2016structural}.

To address the limitations presented by MIL algorithms, this paper proposes a novel support vector machine formulation with a bag-representative selector, called Multiple-Instance Representative Support Vector Machine (MIRSVM). SVMs are known to perform well when data is limited, therefore combining them with a bag-representative selector aims to remedy class imbalance caused by limited positive bag instances, without assuming the bags' internal distributions. The algorithm selects bag-representatives iteratively according to the standard MI assumption, ensuring known information, such as the negative bag labels, are fully utilized during the training process. The optimal separating hyperplane between bags is then found with respect to the bag-representatives using a Gaussian kernel and a quadratic programming solver. The optimal set of representatives is found when they have converged, or stopped changing from one iteration to the next. The algorithm does not assume any distribution of the instances and is not affected by the number of instances within a bag, making it applicable to a variety of contexts. The key contributions of this work include,
\begin{itemize}
\iitem Reformulating the traditional primal L$1$-SVM problem to optimize over bags, rather than instances, ensuring all the information contained within each bag is utilized during training, while defining bag representative selector criteria.

\iitem Deriving the dual multi-instance SVM problem, with the Karush-Kuhn-Tucker necessary and sufficient conditions for optimality. The dual is maximized with respect to the Lagrange multipliers and provides insightful information about the resulting sparse model. The dual formulation is kernelized with a Gaussian radial basis function, which calculates the distances between bag representatives.

\iitem Devising a unique bag-representative selection method that makes no presumptions about the underlying distributions of the instances within each bag, while maintaining the default MI assumption. This approach eliminates the issue of class imbalance caused by techniques such as flattening or subsetting positive instances from each bag. The key feature of MIRSVM is its ability to identify instances (support vectors) within positive and negative bags that highly impact the model.
\end{itemize}

@@

\subsection{Multiple-Instance Learning}\label{subsec:mil}
\begin{sloppypar}
In traditional binary classification problems, the goal is to learn a model that maps input samples to labels, $f: \reals^{m \times d} \rightarrow Y^m \in \set{-1,\,+1}$. In the MIL case, samples are called \textit{bags} and each bag contains one or more input instances and is assigned a single bag-level label. Input instances, $\set{\bm x_1,\, \bm x_2,\, \ldots,\, \bm x_m}$, are grouped into bags with unique identifiers, $\mathcal{B} = \set{\mathcal{B}_1,\, \mathcal{B}_2,\, \ldots,\, \mathcal{B}_n \st \mathcal{B}_I = \set{\bm x_i \st \forall i \in I},\, \forall I \in \set{1, \ldots, n}}$ and assigned a label, $Y_I$. An example MI dataset representation is shown in Figure~\ref{fig:mirep}. Using the training dataset $\mathcal{D} = \{(\mathcal{B}_1,Y_1), \ldots, (\mathcal{B}_n,Y_n)\}$, the goal is to train a classifier that predicts the label of an unseen bag, $f(\mathcal{B}_{n+1}) \rightarrow Y_{n+1}$~\citep{Amores2013}. In order to build a classifier without any knowledge of the individual training instance labels,~\citet{Dietterich1997} proposed the \textit{standard MI} (SMI) hypothesis based on the drug-activity prediction domain, shown in Equation~\eqref{eqn:stas}, which states that a bag is labeled positive if and only if at least one of the instances in the bag is positive, and is labeled negative otherwise.
\end{sloppypar}
\begin{equation}
\centering \label{eqn:stas}
 Y_I = \begin{cases}
			+1 & \text{if }\exists\, y_i = +1,\, \forall i \in I \\
			-1 & \text{otherwise}.
		  \end{cases}
\end{equation}

This implies that individual instance labels $y_i$ exist, but are not known (for positive bags) during training. Equation~\eqref{eqn:stas} can also be rewritten as Equation~\eqref{eqn:stasmax} for simplicity:
\begin{equation}
\centering \label{eqn:stasmax}
Y_I = \text{argmax}_{\forall i \in I}\,\,\, y_i.
\end{equation}

In addition to the SMI assumption, alternative MI assumptions have been proposed to date~\citep{Foulds2010}. A recent review describing the taxonomy of multi-instance classification was presented by~\citet{Amores2013}. The review presents various methods and algorithms used in literature which are categorized based on their approach to handling the MI input space. Instance-based classifiers that fall under the \textit{instance-space paradigm}, aim to separate instances in positive bags from those in negative ones. Bag-level classifiers (\textit{bag-space paradigm}) treat each bag as a whole entity, implicitly extracting information from each bag in order to accurately predict their labels~\citep{qiao2017diversified}. Methods that fall under the \textit{embedded-space paradigm} map the input bags to a single feature vector that explicitly encapsulates all the relevant information contained within each bag.

Instance-based methods that follow the SMI assumption attempt to identify desirable instance properties that make a bag positive. One traditional method in this category is the Axis-Parallel Rectangle (APR)~\citep{Dietterich1997}, which trains a model that assigns a positive label to an instance if it belongs to an axis-parallel rectangle in feature space, and assigns a negative label otherwise. The APR is optimized by maximizing the number of positive bags in the training set containing at least one instance in the APR, while concurrently maximizing the number of negative bags that do not contain any instance in the APR. Another similar method is the Diverse Density (DD)~\citep{Maron1998} framework which is maximized for instances in feature space that are near at least one instance in a positive bag and far from all instances in negative bags. In the Expectation-Maximization Diverse Density (EM-DD) algorithm,~\citet{Zhang2001} propose a similar framework that iteratively maximizes the DD measure.~\citet{Auer2004} present a boosting approach that uses balls centered around positive bags to solve the MI problem called Multi-Instance Optimal Ball (MIOptimalBall). This approach is similar to that of APR and DD, except that~\citet{Auer2004} propose computing optimal balls per positive bags. A major challenge affecting these methods is that the distributions of the positive and negative bags affect their performance. Methods based on the DD metric~\citep{qi2007,Carbonneau2016,Chen2006,Chen2004} assume the positive instances form a cluster, which may not be the case. Alternatively, \citet{Fu2011} models the distribution of negative bags with Gaussian kernels, which can prove difficult when the quantity of data is limited. 
\begin{figure*}
\centering
\begin{tabular}{|c|ccccc|c|}
\hline
$\mathcal{B}$ & $x_1$ & $x_2$ & $x_3$ & \ldots & $x_d$ & $\bm Y$ \\        
\hline
    	  & 0.1 & 0.8 & 2.5 & \ldots & 0.8 &    \\        
    $\mathcal{B}_1$ 	  & 0.2 & 2.0 & 5.5 & \ldots & 3.0 & POS \\
    	  & 0.1 & 0.1   & 4.5 & \ldots & 0.1 &    \\
\hline
    	  & 1.5 & 4.0 & 0.8 & \ldots & 0.1 &    \\
    	  & 0.8 & 0.4 & 2.9 & \ldots & 1.1 &    \\
    $\mathcal{B}_2$	  & 2.3 & 0.2 & 4.0 & \ldots & 5.5 & NEG \\
    	  & 6.7 & 5.0 & 0.1  & \ldots & 0.5 &    \\
    	  & 0.1 & 4.0 & 8.7 & \ldots & 3.3 &    \\
\hline
  \end{tabular}
  \caption{Example multi-instance data representation. The figure shows a sample binary MI dataset with $2$ bags of different sizes and labels, $\mathcal{B}_1$ and $\mathcal{B}_2$. The instances are $d$-dimensional.}%
  \label{fig:mirep}%
\end{figure*}

Some methods in literature~\citep{Ray2005,Viola2005}, such as Simple-MI~\citep{Dong2006}, transform the MI dataset to a traditional instance-level dataset. Simple-MI represents each bag with the mean vector of the instances within it. This approach was evaluated by~\citet{Bunescu2007} and they proposed mapping each bag to a max-min vector, a concatenation of the features with the highest and lowest values. The major disadvantage of these types of approaches is that they assume the distribution the instances in positive bags is positive, when it may not be.

An extension of traditional single-instance $k$-nearest neighbors method (KNN) was proposed by~\citet{Wang2000} to be applied to the bag-level, named CitationKNN. This method uses a distance function between bags in order to determine bag similarities. Not only are the set of closest bags to a single bag  considered, but also how many bags is the single bag closest to. A voting scheme is then used to determine the bag class labels.~\citet{Xu2011} proposed a multi-instance Decision-Based Neural Network (MI-DBNN), a probabilistic variant of the traditional DBNN, which is a neural network with a modular structure~\citep{Herrera2016}.

~\citet{Blockeel2005} introduced the Multi-Instance Tree Inducer (MITI), based on the standard MI assumption, which uses decision trees as a heuristic to solve the MI problem. This approach aims to identify whether an instance within a bag is truly positive and eliminate false positives within the same bag. The disadvantage of this approach stems from removing instances considered as false positives from partially grown trees without updating the existing tree structure.~\citet{Bjerring2011} then enhanced this approach by creating the method Multi-Instance Rule Induction (MIRI). The algorithm aims to eliminate any possibility of a suboptimal split because the tree is discarded and regrown. Other popular methods include MIBoost~\citep{Hall2009}, MIWrapper~\citep{Frank2003}, and Two-Level Classifier (TLC)~\citep{Wang2014}.

For most of the methods described above, implicit or explicit assumptions have been made about the distribution of the data. Selecting a method that is robust for a problem such as MIL can be difficult when little is known about the nature of the data, especially considering the unknown distribution of the instances within bags~\citep{2015-KAIS,2016-INS-MIARM}. The proposed method, MIRSVM, is a general method that uses support vector machines to design a MIL model without making prior assumptions about the data. Classifiers of this type are known to provide better generalization capabilities and performance, as well as sparser models.

\subsection{Support Vector Machines}\label{subsec:svms}
Most classical learning techniques require knowledge of the data distribution to build accurate models. This is a serious restriction because, in most cases, the data distribution is unknown~\citep{Kecman2001}. SVMs represent learning techniques that have been introduced under the \textit{structural risk minimization} (SRM) framework and VC theory. Rather than optimizing over L1 or L2 norms and classification error, SVMs perform SRM~\citep{Shalev2014}, minimizing the expected probability of classification error, resulting in a generalized model without making assumptions about the data distribution~\citep{Cortes1995}.

SVMs are a particularly useful for learning linear predictors in high dimensional feature spaces, which is a computationally complex problem~\citep{wang2016hierarchical}. In the context of classification, this problem is approached by searching for the optimal \textit{maximal margin} of separability between classes~\citep{carrasco2015multi,fan2017mreklm}. Since most real-world datasets are non-linearly separable, ~\citet{Cortes1995} introduced the \textit{Soft-Margin L$1$-SVM}:
\begin{subequations} 
\label{eq:primalsvm}
\begin{alignat}{2}
\min\limits_{\bm w,b,\bm\xi}& {\,\,} \frac{1}{2}||\bm{w}||^2 + \frac{C}{m}\sum_{i=1}^m \xi_i \tag{\ref{eq:primalsvm}} \\ 
\text{s.t.}& {\,\,} y_i(\langle\bm{w},\bm{x}_i\rangle + b) \geq 1 - \xi_i,\,\, \forall i \in \{1,\ldots,m\}  \\
{} & {\,\,} \xi_i \geq 0,\,\, \forall i \in \{1,\ldots,m\},
\end{alignat}
\end{subequations} 
where $\bm w \in \reals^d$ is a $d$-dimensional weight vector, $b \in \reals$ is a bias term, and $C \in \reals$ is the penalty parameter that controls the trade-off between margin maximization and classification error minimization. The slack variable $\bm \xi \in \reals^m$ allows for optimizing over sample errors. However, if the training dataset is not linearly separable, the classifier may not have a good generalization capability~\citep{Nocedal2006}. Generalization and linear separability can be enhanced by mapping the original input space to a higher dimensional dot-product space by using a kernel function shown in Equation~\eqref{eqn:kerneltrick}:
\begin{equation}
\spa{K}\left(\bm{x}_i,\bm{x}_j\right) = \langle \phi\left(\bm{x}_i\right),\,\phi\left(\bm{x}_j\right)\rangle,
\label{eqn:kerneltrick}
\end{equation}
where $\phi\left(\cdot\right)$ represents a function mapping from the original feature space to a higher dimensional space. This kernel mapping is helpful when solving the dual SVM problem shown in Equation~\eqref{eq:dualsvm}.
\begin{subequations}
\label{eq:dualsvm}
\begin{alignat}{2}
\max\limits_{\bm \alpha} &\,\,\,\, \frac{1}{2}\sum_{i,j=1}^m{\alpha_i \alpha_jy_iy_j\spa{K}\left(\bm{x}_i,\bm{x}_j\right)} - \sum_{i=1}^m \alpha_i \tag{\ref{eq:dualsvm}} \\ 
\text{s.t.} &\,\,\,\, \sum_{i=1}^m \alpha_iy_i = 0,  \\
&\,\,\,\, 0 \leq \alpha_i \leq \frac{C}{m},\,\, \forall i \in \{1,\ldots,m\}. 
\end{alignat}
\end{subequations}

@@

\citet{Andrews2002} proposed a mixed-integer quadratic program that solves the MI problem at an instance-level, using a support vector machine, named MISVM, that can be solved heuristically. Rather than maximizing the margin of separability between instances of different classes, this instance-based method maximizes the margin between bags. It tries to identify the key instance from each positive bag that makes the bag positive by assuming it has the highest margin value~\citep{Herrera2016}. Instances from positive bags are selected as bag-representatives, and the algorithm iteratively creates a classifier that separates those representatives from all instances from the negative bags. Using bag-representatives from one class and all instances from the other is an example of an approach that combines rules from the SMI assumption and the collective assumption. This approach was used in the context of active learning.~\citet{Wang20143106} proposed a multi-criteria decision making system that measures the significance of unlabeled instances and selects the best instance iteratively using MISVM. A disadvantage of the approach MISVM takes, stems from the assumption that all instances within positive bags are also positive, which is an implicit step in the initialization of MISVM.~\citet{Andrews2002} also proposed a second mixed-integer instance-level approach, named mi-SVM, which does not discard the negative instances of the positive bags. It rather tries to identify the instances within positive bags that are negative and utilize them in the construction of the negative margin. The main disadvantage of these approaches is that they create an imbalanced class problem that favors the negative class, resulting in a biased classifier.

@@

One of the first bag-level approaches to the multi-instance SVM problem was proposed by \citet{Smola2002}, who defined a bag-level kernel, which determines the similarity between two bags in a higher dimensional space. Using such a kernel, the margin can be optimized over bags without any modification to the SVM problem~\citep{Herrera2016}.~\citet{Blaschko2006} propose conformal kernels which manipulate each attribute's dimension based on its importance.

@@

\section{Multiple-Instance Representative SVM}\label{sec:contribution}
MIRSVM is based on the idea of selecting representative instances from both positive and negative bags which are used to find an unbiased, optimal separating hyperplane. A representative is iteratively chosen from each bag, and a new hyperplane is formed according to the representatives until they converge. Based on the SMI hypothesis, only one instance in a bag is required to be positive for the bag to adopt a positive label. Due to the unknown distribution of instances within positive bags, MIRSVM is designed to give preference to negative bags during training, because their distribution is known, i.e. all instances are guaranteed to be negative. This is evident during the representative selection process, by taking the index of the maximum output value within each bag based on the current hyperplane using the following rule, $s_I = \text{argmax}_{i \in I} (\langle \bm w,\, \bm x_i \rangle + b),\, \forall I \in \set{1,\ldots,n}$. In other words, the most positive instance is chosen from each positive bag and the least negative instance is chosen from each negative bag (instances with the largest output value based on the current hyperplane), pushing the decision boundary  towards the positive bags. Equation~\eqref{eq:primalmirsvm} presents the primal MIRSVM optimization problem:
\begin{subequations} 
\label{eq:primalmirsvm}
\begin{alignat}{2}
\min\limits_{\bm w, b,\bm\xi}& {\,\,} \frac{1}{2}||\bm{w}||^2 + \frac{C}{n}\sum_I \xi_I, \tag{\ref{eq:primalmirsvm}}\\ 
\text{s.t.}& {\,\,} Y_I(\langle\bm{w},\bm x_{s_I}\rangle + b) \geq 1 - \xi_I,\, \forall I \in \set{1,\ldots,n}  \label{eq:primalmirsvm1}\\
{} & {\,\,} \xi_I \geq 0,\, \forall I \in \set{1,\ldots,n}, 
\end{alignat}
\end{subequations} 
where $\bm S_I$ is a set of the bag representatives' indices and $\bm x_{s_I}$ is the instance representative of bag $\mathcal{B}_I$. Note the variables in MIRSVMs formulation are the similar to those of the traditional SVM, except they are now representing each bag as an instance. Solving the optimization problem given in Equation~\eqref{eq:primalmirsvm} using a quadratic programming solver is a computationally expensive task due to the number of constraints, which scales by the number of bags $n$, as well as the calculation of the inner product between two $d$-dimensional vectors in constraint~\eqref{eq:primalmirsvm1}. The proposed solution for these problems was deriving and solving the dual of the optimization problem given by Equation~\eqref{eq:primalmirsvm}.

The dual can be formed by taking the Lagrangian of~\eqref{eq:primalmirsvm}, given by Equation~\eqref{eqn:svmlagrangian}, where $\bm \alpha$ and $\bm \beta$ are the non-negative Lagrange multipliers.
\begin{equation}
\centering \label{eqn:svmlagrangian}
\resizebox{.9\textwidth}{!}{$
\begin{aligned}
\mathcal{L}\left(\bm{w},b,\bm{\xi},\bm{\alpha},\bm{\beta}\right) = \frac{1}{2}\sum_{j=1}^d w_j^2  + \frac{C}{n}\sum_I \xi_I - \sum_I{\beta_I \xi_I} - \sum_I{\alpha_I\left(Y_I\left(\sum_{j=1}^d w_j x_{s_Ij} + b\right) -1 + \xi_I\right)}
\end{aligned}$} 
\end{equation}

>>>>

\section{Introduction}\label{intro}
In traditional classification, learning algorithms attempt to correctly label unknown samples by finding patterns between training samples and their class label. \textit{Multi-instance} (MI) learning is a generalization of \textit{supervised learning} that has been recently been gaining interest because of its applicability to many real-world problems such as webmining and text categorization {\citep{Andrews2002}}, image classification and retreival \citep{qi2007,cao2017weakly}, human action recognition \citep{Yi2016}, medical diagnosis and imaging {\citep{Herrera2016}}, and bioinformatics {\citep{Dietterich1997}} (drug activity prediction and protein identification).

@@

Support vector machines represent a set of supervised, linear and nonlinear, classification and regression methods that have theoretical foundations on Vapnik-Chervonenkis (VC) theory~\citep{Huang2006,Schoelkopf2002}. SVM models are similar to other machine learning techniques, but research has shown that they usually outperform them in terms of computational efficiency, scalability, and robustness against outliers~\citep{Kecman2016,Melki2016}, making them a useful tool for various real-world applications~\citep{chen2016structural}.

To address the limitations presented by MIL algorithms, this paper proposes a novel support vector machine formulation with a bag-representative selector, called Multiple-Instance Representative Support Vector Machine (MIRSVM). SVMs are known to perform well when data is limited, therefore combining them with a novel bag-representative selector aims to remedy class imbalance caused by limited positive bag instances, without assuming the bags' internal distributions. The algorithm selects bag-representatives iteratively according to the standard MI assumption, ensuring known information, such as the negative bag labels, are fully utilized during the training process. The optimal separating hyperplane between bags is then found with respect to the bag-representatives using a Gaussian kernel and a quadratic programming solver. The optimal set of representatives is found when they have converged, or stopped changing from one iteration to the next. The algorithm does not assume any distribution of the instances and is not affected by the number of instances within a bag, making it applicable to a variety of contexts. The key contributions of this work include,
\begin{itemize}
\iitem Reformulating the traditional primal L$1$-SVM problem to optimize over bags, rather than instances, ensuring all the information contained within each bag is utilized during training, while defining bag representative selector criteria.

\iitem Deriving the dual multi-instance SVM problem, with the Karush-Kuhn-Tucker necessary and sufficient conditions for optimality. The dual is maximized with respect to the Lagrange multipliers and provides insightful information about the resulting sparse model. The dual formulation is kernelized with a Gaussian radial basis function, which calculates the distances between bag representatives.
\newpage
\iitem Devising a unique bag-representative selection method that makes no presumptions about the underlying distributions of the instances within each bag, while maintaining the default MI assumption. This approach eliminates the issue of class imbalance caused by techniques such as flattening or subsetting positive instances from each bag. The key feature of MIRSVM is its ability to identify instances (support vectors) within positive and negative bags that highly impact the model.
\end{itemize}

@@

\subsection{Multiple-Instance Learning}\label{subsec:mil}
\begin{sloppypar}
In traditional binary classification problems, the goal is to learn a model that maps input samples to labels, $f: \reals^{m \times d} \rightarrow Y^m \in \set{-1,\,+1}$. In the MIL case, samples are called \textit{bags} and each bag contains one or more input instances and is assigned a single bag-level label. Input instances, $\set{\bm x_1,\, \bm x_2,\, \ldots,\, \bm x_m}$, are grouped into bags with unique identifiers, $\mathcal{B} = \set{\mathcal{B}_1,\, \mathcal{B}_2,\, \ldots,\, \mathcal{B}_n \st \mathcal{B}_I = \set{\bm x_i \st \forall i \in I},\, \forall I \in \set{1, \ldots, n}}$ and assigned a label, $Y_I$. An example MI dataset representation is shown in Figure~\ref{fig:mirep}. Using the training dataset $\mathcal{D} = \{(\mathcal{B}_1,Y_1), \ldots, (\mathcal{B}_n,Y_n)\}$, the goal is to train a classifier that predicts the label of an unseen bag, $f(\mathcal{B}_{n+1}) \rightarrow Y_{n+1}$. In order to build a classifier without any knowledge of the individual training instance labels,~\citet{Dietterich1997} proposed the \textit{standard MI} (SMI) hypothesis based on the drug-activity prediction domain, shown in Equation~\eqref{eqn:stas}, which states that a bag is labeled positive if and only if at least one of the instances in the bag is positive, and is labeled negative otherwise.
\end{sloppypar}
\begin{equation}
\centering \label{eqn:stas}
 Y_I = \begin{cases}
			+1 & \text{if }\exists\, y_i = +1,\, \forall i \in I \\
			-1 & \text{otherwise}.
		  \end{cases}
\end{equation}

This implies that individual instance labels $y_i$ exist, but are not known (for positive bags) during training. Equation~\eqref{eqn:stas} can also be rewritten as $Y_I = \text{argmax}_{\forall i \in I}\,\,\, y_i$, for simplicity.

In addition to the SMI assumption, alternative MI assumptions have been proposed to date~\citep{Foulds2010}. A recent review describing the taxonomy of multi-instance classification was presented by \citet{Amores2013}. Classifiers that fall under the \textit{instance-space paradigm} aim to separate instances in positive bags from those in negative ones. Bag-level classifiers (\textit{bag-space paradigm}) treat each bag as a whole entity, implicitly extracting information from each bag in order to accurately predict their labels~\citep{qiao2017diversified}. Methods that fall under the \textit{embedded-space paradigm} map the input bags to a single feature vector that explicitly encapsulates all the relevant information contained within each bag.

Instance-based methods under the SMI assumption attempt to identify desirable instance properties that make a bag positive. One traditional method is the Axis-Parallel Rectangle (APR) \citep{Dietterich1997}, which assigns a positive label to an instance if it belongs to an APR. The Diverse Density (DD){~\citep{Maron1998}} measure is maximized for instances that are near at least one instance in a positive bag and far from all instances in negative bags.~\citet{Auer2004} present a boosting approach that uses balls centered around positive bags, called Multi-Instance Optimal Ball (MIOptimalBall). A major challenge of these methods is that the distributions of positive and negative bags affect their performance~\citep{Chen2006,Chen2004}. Alternatively \citet{Fu2011} models the distribution of negative bags with Gaussian kernels, which can prove difficult when the quantity of data is limited.

Some methods in literature~\citep{Ray2005,Viola2005}, such as Simple-MI~\citep{Dong2006}, transform the MI dataset to a traditional instance-level dataset. This approach was evaluated by~\citet{Bunescu2007} and they proposed mapping each bag to a max-min vector. The disadvantage of these approaches is that they assume the labels of all instances within positive bags are positive, when this may not be the case.
\begin{figure*}
\centering
\small
\begin{tabular}{|c|ccccc|c|}
\hline
$\mathcal{B}$ & $x_1$ & $x_2$ & $x_3$ & \ldots & $x_d$ & $\bm Y$ \\        
\hline
    	  & 0.1 & 0.8 & 2.5 & \ldots & 0.8 &    \\        
    $\mathcal{B}_1$ 	  & 0.2 & 2.0 & 5.5 & \ldots & 3.0 & POS \\
    	  & 0.1 & 0.1   & 4.5 & \ldots & 0.1 &    \\
\hline
    	  & 1.5 & 4.0 & 0.8 & \ldots & 0.1 &    \\
    	  & 0.8 & 0.4 & 2.9 & \ldots & 1.1 &    \\
    $\mathcal{B}_2$	  & 2.3 & 0.2 & 4.0 & \ldots & 5.5 & NEG \\
    	  & 6.7 & 5.0 & 0.1  & \ldots & 0.5 &    \\
    	  & 0.1 & 4.0 & 8.7 & \ldots & 3.3 &    \\
\hline
  \end{tabular}
  \caption{Example multi-instance data representation. The figure shows a sample binary MI dataset with $2$ bags of different sizes and labels, $\mathcal{B}_1$ and $\mathcal{B}_2$. The instances are $d$-dimensional.}%
  \label{fig:mirep}%
\end{figure*}

An extension of traditional single-instance $k$-nearest neighbors method (KNN) was proposed by~\citet{Wang2000} to be applied to the bag-level, named CitationKNN. This method uses a distance function between bags in order to determine bag similarities and a voting scheme to determine the bag class labels.~\citet{Xu2011} proposed a multi-instance Decision-Based Neural Network (MI-DBNN), which is a neural network with a modular structure{~\citep{Herrera2016}}.

~\citet{Blockeel2005} introduced the Multi-Instance Tree Inducer (MITI), which uses decision trees as a heuristic to solve the MI problem. MITI aims to identify whether an instance within a bag is truly positive and eliminate false positives within the same bag. The disadvantage of this approach stems from removing instances considered as false positives from partially grown trees without updating the existing tree structure.~\citet{Bjerring2011} then enhanced this approach by creating the method Multi-Instance Rule Induction (MIRI), which eliminates any possibility of a suboptimal split because the tree is discarded and regrown. Other popular methods include MIBoost~\citep{Hall2009}, MIWrapper~\citep{Frank2003}, and Two-Level Classifier (TLC)~\citep{Wang2014}.

For most of the methods described above, implicit or explicit assumptions have been made about the distribution of the data~\citep{2017-KNOSYS-MIMV}. Selecting a method that is robust for a problem such as MIL can be difficult when little is known about the nature of the data, especially considering the unknown distribution of the instances within bags~\citep{2015-KAIS,2016-INS-MIARM}. MIRSVM, is an iterative method that uses SVMs to design a MIL model without making prior assumptions about the data. Classifiers of this type are known to provide better generalization capabilities, performance, and sparser models.

\subsection{Support Vector Machines}\label{subsec:svms}
Most classical learning techniques require knowledge of the data distribution to build accurate models, which is a serious restriction because, in most cases, the distribution is unknown~\citep{Kecman2001}. SVMs represent learning techniques that have been introduced under the \textit{structural risk minimization} (SRM) framework and VC theory. Rather than optimizing over L1 or L2 norms and classification error, SVMs perform SRM~\citep{Shalev2014}, minimizing the expected probability of classification error, resulting in a generalized model without making assumptions about the data distribution~\citep{Cortes1995}.

SVMs are a particularly useful for learning linear predictors in high dimensional feature spaces, which is a computationally complex problem~\citep{wang2016hierarchical}. In the context of classification, this problem is approached by searching for the optimal \textit{maximal margin} of separability between classes~\citep{carrasco2015multi,fan2017mreklm}. Since most real-world datasets are non-linearly separable, ~\citet{Cortes1995} introduced the \textit{Soft-Margin L$1$-SVM}. However, if the training dataset is not linearly separable, the classifier may not have a good generalization capability~\citep{Nocedal2006}. Generalization and linear separability can be enhanced by mapping the original input space to a higher dimensional dot-product space by using a kernel function shown in Equation~\eqref{eqn:kerneltrick}:
\begin{equation}
\spa{K}\left(\bm{x}_i,\bm{x}_j\right) = \langle \phi\left(\bm{x}_i\right),\,\phi\left(\bm{x}_j\right)\rangle,
\label{eqn:kerneltrick}
\end{equation}
where $\phi\left(\cdot\right)$ represents a function mapping from the original feature space to a higher dimensional space. This kernel mapping is helpful when solving the dual SVM problem shown in Equation~\eqref{eq:dualsvm}.
\begin{subequations}
\label{eq:dualsvm}
\begin{alignat}{2}
\max\limits_{\bm \alpha} &\,\,\,\, \frac{1}{2}\sum_{i,j=1}^m{\alpha_i \alpha_jy_iy_j\spa{K}\left(\bm{x}_i,\bm{x}_j\right)} - \sum_{i=1}^m \alpha_i \tag{\ref{eq:dualsvm}} \\ 
\text{s.t.} &\,\,\,\, \sum_{i=1}^m \alpha_iy_i = 0,  \\
&\,\,\,\, 0 \leq \alpha_i \leq \frac{C}{m},\,\, \forall i \in \{1,\ldots,m\}, 
\end{alignat}
\end{subequations}

where $C \in \reals$ is the penalty parameter that controls the trade-off between margin maximization and classification error minimization.

@@

\citet{Andrews2002} proposed a mixed-integer quadratic program that solves the MI problem at an instance-level, using a support vector machine, named MISVM, that can be solved heuristically. Rather than maximizing the margin of separability between instances of different classes, this instance-based method maximizes the margin between bags. It tries to identify the key instance from each positive bag that makes the bag positive by assuming it has the highest margin value~\citep{Herrera2016}. Using bag-representatives from the positive class class and all instances from the other is an example of an approach that combines rules from the SMI assumption and the collective assumption.~\citet{Wang20143106} proposed a multi-criteria decision making system that measures the significance of unlabeled instances and selects the best instance iteratively using MISVM. A disadvantage of this approach stems from the assumption that all instances within positive bags are also positive, which is an implicit step in the initialization of MISVM.~\citet{Andrews2002} also proposed a second mixed-integer instance-level approach, named mi-SVM, which tries to identify the instances within positive bags that are negative and utilize them in the construction of the negative margin. The main disadvantage of these approaches is that they create an imbalanced class problem that favors the negative class, resulting in a biased classifier.

@@

One of the first bag-level approaches to the multi-instance SVM problem was proposed by \citet{Smola2002}, who defined a bag-level kernel, which determines the similarity between two bags in a higher dimensional space.~\citet{Blaschko2006} propose conformal kernels which manipulate each attribute's dimension based on its importance.

@@

\section{Multiple-Instance Representative SVM}\label{sec:contribution}
MIRSVM is based on the idea of selecting representative instances from both positive and negative bags which are used to find an unbiased, optimal separating hyperplane. A representative is iteratively chosen from each bag, and a new hyperplane is formed according to the representatives until they converge. Based on the SMI hypothesis, only one instance in a bag is required to be positive for the bag to adopt a positive label. Due to the unknown distribution of instances within positive bags, MIRSVM is designed to give preference to negative bags during training, because their distribution is known, i.e. all instances are guaranteed to be negative. This is evident during the representative selection process, by taking the index of the maximum output value within each bag based on the current hyperplane using the following rule:
\begin{equation}
\label{eq:SIPrimal}
\centering
s_I = \text{argmax}_{i \in I} (\langle \bm w,\, \bm x_i \rangle + b),\, \forall I \in \set{1,\ldots,n},
\end{equation}
where $\bm w \in \reals^d$ is a $d$-dimensional weight vector, $b \in \reals$ is a bias term. In other words, the most positive instance is chosen from each positive bag and the least negative instance is chosen from each negative bag (instances with the largest output value based on the current hyperplane), pushing the decision boundary  towards the positive bags. Equation~\eqref{eq:primalmirsvm} presents the primal MIRSVM optimization problem:
\begin{subequations} 
\label{eq:primalmirsvm}
\begin{alignat}{2}
\min\limits_{\bm w, b,\bm\xi}& {\,\,} \frac{1}{2}||\bm{w}||^2 + \frac{C}{n}\sum_I \xi_I, \tag{\ref{eq:primalmirsvm}}\\ 
\text{s.t.}& {\,\,} Y_I(\langle\bm{w},\bm x_{s_I}\rangle + b) \geq 1 - \xi_I,\, \forall I \in \set{1,\ldots,n}  \label{eq:primalmirsvm1}\\
{} & {\,\,} \xi_I \geq 0,\, \forall I \in \set{1,\ldots,n}, 
\end{alignat}
\end{subequations} 
where $\bm S_I$ is a set of the bag representatives' indices, $\bm x_{s_I}$ is the instance representative of bag $\mathcal{B}_I$, and the slack variable $\bm \xi \in \reals^m$ allows for optimizing over bag errors. Note the variables in MIRSVMs formulation are the similar to those of the traditional SVM, except they are now representing each bag as an instance. This is one of the benefits of the MIRSVM formulation, whereby using a bag-representative selector, traditional SVM solvers can be used for new data representations such the multi-instance data.

Solving the optimization problem given in Equation~\eqref{eq:primalmirsvm} using a quadratic programming solver is a computationally expensive task due to the number of constraints, which scales by the number of bags $n$, as well as the calculation of the inner product between two $d$-dimensional vectors in constraint~\eqref{eq:primalmirsvm1}. The proposed solution for these problems was deriving and solving the dual of the optimization problem given by Equation~\eqref{eq:primalmirsvm}.
